Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Add missing apostrophes

Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, 
and apparently accepts it, as evidenced by his use\footnote{
  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, 
  respectively, Lemma 4 and Theorem 4.
} 
of A4 and A5, as well as varying domain semantics, in his new
emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
axiom H:A12. Surprisingly, the computer-assisted formalization of
\AOEH\ shows that A4 and A5 are still superfluous. Moreover, 
A4 and A5 are independent\footnote{
  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff 
  there are models of $S$ where $A$ is true and other models of 
  $S$ where $A$ if false. 
} 
of the other axioms and definitions. 
Therefore, A4 and A5 are not redundant, despite their superfluousness.

Although, Hájek did not notice the superfluousness of A4 and A5 in his
\AOEH, he did describe yet another emendation (his \AOEHH, cf.
Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no
claim is made w.r.t. to their redundancy), if A3 is replaced by a
stronger axiom (H:A3) additionally stating that the property of actual
existence is positive when it comes to God-like beings
\cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
not only superflous, but also redundant. For A5, no conclusive results
were achieved; neither a proof nor a countermodel could be
automatically generated. Surprisingly, a countermodel for the weaker
A3 was succesfully generated. This is somewhat unsatisfactory (for
theistic goals), because it shows that \AOEHH\ does not entail the
positiveness of being God-like.

Nevertheless, \AOEHH\ is explicitly regarded by
\citet[p.~12]{Hajek2002} as just an intermediary step towards a more
natural theory, based on a more sophisticated notion of positiveness.
That is his final emendation (\AOEHHH, cf. \ref{apx:Hajek3}), which
restores A3 and does use A4 and A5, albeit in a modified form (i.e.
H:A4 and H:A5). The formalization of \AOEHHH\ shows that H:A4 is
independent. The old A5 is independent as well, and both H:A4 and H:A5
are superfluous, but no conclusive results were achieved regarding
independence or redundancy of H:A5.

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant, but A5 becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

@@

% ToDo: \\
% 3 - Fix Hajek's proofs in the appendices. (they are just copy-pasted from Anderson's so far)
% 4 - Check apostrophes in axiom names that depend on different definitions.
% 5 - Add subscripts to alternative definitions (e.g. G_A)

>>>>

Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, 
and apparently accepts it, as evidenced by his use\footnote{
  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, 
  respectively, Lemma 4 and Theorem 4.
} 
of A4 and A5', as well as varying domain semantics, in his new
emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
axiom H:A12. Surprisingly, the computer-assisted formalization of
\AOEH\ shows that A4 and A5' are still superfluous. Moreover, 
A4 and A5' are independent\footnote{
  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff 
  there are models of $S$ where $A$ is true and other models of 
  $S$ where $A$ if false. 
} 
of the other axioms and definitions. 
Therefore, A4 and A5' are not redundant, despite their superfluousness.

Although Hájek did not notice the superfluousness of A4 and A5' in his
\AOEH, he did describe yet another emendation (his \AOEHH, cf.
Appendix \ref{apx:Hajek2}) where A4 and A5' are superfluous (though no
claim is made w.r.t. to their redundancy), if A3' is replaced by a
stronger axiom (H:A3) additionally stating that the property of actual
existence is positive when it comes to God-like beings
\cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
not only superflous, but also redundant. For A5', no conclusive results
were achieved; neither a proof nor a countermodel could be
automatically generated. Surprisingly, a countermodel for the weaker
A3' was succesfully generated. This is somewhat unsatisfactory (for
theistic goals), because it shows that \AOEHH\ does not entail the
positiveness of being God-like.

Nevertheless, \AOEHH\ is explicitly regarded by
\citet[p.~12]{Hajek2002} as just an intermediary step towards a more
natural theory, based on a more sophisticated notion of positiveness.
That is his final emendation (\AOEHHH, cf. \ref{apx:Hajek3}), which
restores A3' and does use A4 and A5', albeit in a modified form (i.e.
H:A4 and H:A5). The formalization of \AOEHHH\ shows that H:A4 is
independent. The old A5' is independent as well, and both H:A4 and H:A5
are superfluous, but no conclusive results were achieved regarding
independence or redundancy of H:A5.

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant, but A5' becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

@@

% ToDo: \\
